---
title: "Cox PH Models"
author: "Landi Luo"
date: "12/7/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages:
```{r, message=FALSE}
if (!require("pacman"))  
  install.packages("pacman", repos = "http://cran.us.r-project.org/")
p_load("tidyverse", "survival", "kableExtra")
```

Import data:
```{r}
breast <- readRDS(file = "breast_final.rds")
```

## Cox Model

Using the Breslow method of handling ties, we fit a Cox proportional hazards model to the data including all 13 covariates: race, sex, stage, breast subtype, age dx, age, marital status, benign tumor count, malignant tumor count, primary site, pr status, er status, insurance status.

```{r}
fit <- coxph(Surv(SRV_TIME_MON, delta) ~ factor(stage) + factor(RAC_RECY) + 
             factor(SEX) + factor(BRST_SUB) + AGE_DX + Age + factor(MAR_STAT) + 
             MALIGCOUNT + BENBORDCOUNT + factor(PRIMSITE) + factor(ERSTATUS) +
             factor(PRSTATUS) + factor(INSREC_PUB), 
             data = breast, ties = "breslow" )

#summary(fit)

```

## ANOVA Table

We constructed an Analysis of Variance table to summarize estimates of the risk coefficients and the results of the one degree of freedom tests for each covariate in the model:
```{r}
anova_table <- data.frame(summary(fit)$coefficients)
kable(anova_table, "latex", booktabs = T,
      col.names = c("Coefficient", "Exp. Coeff.", "Std. Error", "Z-Score", "P-Value")) %>%
  kable_styling(latex_options = c("striped", "HOLD_position"))
```

