---
title: "group project"
author: "Lumi Huang"
date: "11/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(dplyr)
```

```{r}
breast <- read_csv(file = "./seer_breast.csv")
#breast <- as.data.frame.list(breast)
dim(breast)
breast_ca <- breast %>% 
  filter(REG == '0000001541'| REG == '0000001501'| REG == '0000001531' | REG == '0000001535') %>%
  filter(YEAR_DX == 2014)
dim(breast_ca)
```

Variables of interest:
STAT_REC = Vital status recode (study cutoff used)  censoring indicators!
VSRTSADX  = SEER cause of death classification    censoring indicators?
ODTHCLASS = SEER other cause of death classification
SRV_TIME_MON = Survival months survival time? not sure the difference between the 2
SRV_TIME_MON_FLAG = Survival months flag
AGE_DX = Age at diagnosis
MDXRECMP = Month of diagnosis                  
YEAR_DX = Year of diagnosis
MALIGCOUNT = Total number of in situ/malignant tumors for patient       
BENBORDCOUNT = Total number of benign/borderline tumors for patient
RACE1V = Race/ethnicity
SEX = Sex
EOD10_PN = positive lymph nodes examined
CODPUB = Cause of death to SEER site recode
```{r}
breast_data <- breast_ca %>% 
  select(STAT_REC, VSRTSADX, ODTHCLASS, SRV_TIME_MON, SRV_TIME_MON_FLAG,
         AGE_DX, MDXRECMP, YEAR_DX, MALIGCOUNT, BENBORDCOUNT, RACE1V, SEX,
         EOD10_PN, CODPUB)
```