---
title: "group project"
author: "Lumi Huang"
date: "11/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clean the seer_breast dataset

```{r}
library(tidyverse)
library(dplyr)
```

Variables of interest:
STAT_REC = Vital status recode (study cutoff used)  censoring indicators!
VSRTSADX  = SEER cause of death classification    censoring indicators
SRV_TIME_MON = Survival months survival time
AGE_DX = Age at diagnosis
MALIGCOUNT = Total number of in situ/malignant tumors for patient       
BENBORDCOUNT = Total number of benign/borderline tumors for patient
RAC_RECY = race
SEX = Sex
BRST_SUB = BREAST SUBTYPE
EOD10_PN = positive lymph nodes examined
MAR_STAT = marital status at dx
PRIMSITE = Primary Site
     C50.0    Nipple
     C50.1    Central portion of breast
     C50.2    Upper-inner quadrant of breast
     C50.3    Lower-inner quadrant of breast
     C50.4    Upper-outer quadrant of breast
     C50.5    Lower-outer quadrant of breast
     C50.6    Axillary tail of breast
     C50.8    Overlapping lesion of breast
     C50.9    Breast, NOS

```{r}
breast <- read_csv(file = "./seer_breast.csv")
dim(breast)
breast_ca <- breast %>% 
  filter(REG == '0000001541')
dim(breast_ca)

#drop region
breast_ca <- breast_ca %>% select(-REG)
dim(breast_ca)

#filter the data so that we only look at those who die from breast cancer and those who survive till the end of the study
breast_ca <- breast_ca %>%
  filter((VSRTSADX == 0 & STAT_REC == 1) | VSRTSADX == 1)

#create new death indicator delta: delta = 0 alive till the end of the study, 1 = die from breast cancer
breast_ca$delta <- ifelse(breast_ca$VSRTSADX == 1, 1, 0)

#create new variable stage: stage 0 - 4
breast_ca <- breast_ca %>%
  mutate(stage = recode(DAJCC7STG, "120" = 1, "000" = 0, "320" = 2, "520" = 3, "999" = 9, "330" = 2,
                        "700" = 4, "150" = 1, "540" = 3, "530" = 3, "888" = 9, "510" = 3))
#drop stage = NA
breast_ca <- breast_ca %>% filter(stage != 9)

#drop race = unknown
breast_ca <- breast_ca %>% filter(RAC_RECY != 9)

#drop brest subtype = unknown (5)
breast_ca <- breast_ca %>% filter(BRST_SUB != 5)

#drop unknown survival time months (9999)
breast_ca <- breast_ca %>% filter(SRV_TIME_MON != 9999)

#drop age unknown (999)
breast_ca <- breast_ca %>% filter(AGE_DX != 999)

#drop unknown MAR_STAT = 9
breast_ca <- breast_ca %>% filter(MAR_STAT != 9)

#drop unknown MALIGCOUNT = 99 BENBORDCOUNT = 99
breast_ca <- breast_ca %>% filter(MALIGCOUNT != 99)
breast_ca <- breast_ca %>% filter(BENBORDCOUNT != 99)

#transform character variables into numeric
breast_ca$MALIGCOUNT <- as.numeric(breast_ca$MALIGCOUNT)
breast_ca$BENBORDCOUNT <- as.numeric(breast_ca$BENBORDCOUNT)
breast_ca$SRV_TIME_MON <- as.numeric(breast_ca$SRV_TIME_MON)

breast_ca$AGE_DX <- as.numeric(breast_ca$AGE_DX)
#convert YR_BRTH to age
breast_ca$Age <- 2015 - breast_ca$YR_BRTH


breast_ca$PRIMSITE
#convert PRIMSITE to numeric
breast_ca <- breast_ca %>%
  mutate(PRIMSITE = recode(PRIMSITE, "C500" = 0, "C501" = 1, "C502" = 2, "C503" = 3, "C504" = 4,
                           "C505" = 5, "C506" = 6, "C508" = 7, "C509" = 8))

breast_ca <- breast_ca %>%
  filter(ERSTATUS != 3 & ERSTATUS != 4)
PRSTATUS
INSREC_PUB



```

save an RDS file with the following covariates (for easier access later):

- stage, delta, RAC_RECY, SEX, SRV_TIME_MON, BRST_SUB, AGE_DX, YR_BRTH, MAR_STAT, MALIGCOUNT, BENBORDCOUNT
```{r}
setwd(".")
source("Rfunction.R")

single.cox(breast_ca$SRV_TIME_MON, breast_ca$delta, as.data.frame(breast_ca[, c("MALIGCOUNT", "RAC_RECY")]), c("N","C"))
breast_final <- breast_ca %>%
  select(stage, delta, RAC_RECY, SEX, SRV_TIME_MON, BRST_SUB, AGE_DX, YR_BRTH,
         MAR_STAT, MALIGCOUNT, BENBORDCOUNT)
saveRDS(breast_final, file = "breast_final.rds")  
```


