---
title: "M215 Project Analyses"
author: "Landi Luo"
date: "12/6/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(survival)
library(ncvreg)
```

Import data
```{r}
setwd(".")
breast_ca <- readRDS(file = "breast_final.rds")
```

```{r}
#convert YR_BRTH to age
breast_ca$Age <- 2014 - breast_ca$YR_BRTH
breast_ca <- breast_ca %>% select(-YR_BRTH)
```
## Variable selection

variables:

- stage: cancer stage
- delta: death indicator
- SRV_TIME_MON = Survival months survival time
- SEX = Sex
- RAC_RECY = race
- MAR_STAT = marital status at dx
- BRST_SUB = BREAST SUBTYPE
- AGE_DX = age at diagnosis
- YR_BRTH = year of birth
- MALIGCOUNT = Total number of in situ/malignant tumors for patient       
- BENBORDCOUNT = Total number of benign/borderline tumors for patient
- YR_BRTH = Age (year of end of study - year of birth)

```{r}
source("Rfunction.R")
```

The design matrix of predictor values
```{r}
X <- breast_ca %>% 
  select(-SRV_TIME_MON, -delta)
```

Lasso
```{r}
lasso.fit <- ncvsurv(X, Surv(SRV_TIME_MON, delta), penalty = "lasso", nlambda = 25) 
lasso.bic <- AIC(lasso.fit, k = log(n))
lasso.est <- lasso.fit$beta[, which.min(lasso.bic)]
wt <- 1 / abs(coxph(Surv(SRV_TIME_MON, delta) ~ X)$coefficient)
alasso.fit <- ncvsurv(X, Surv(SRV_TIME_MON, delta), penalty = "lasso", nlambda = 25, penalty.factor = wt)
alasso.bic <- AIC(alasso.fit, k = log(n))
alasso.est <- alasso.fit$beta[, which.min(alasso.bic)]
```