---
title: "Survival Final Project Landi RMD Code"
author: "Landi Luo"
date: "12/6/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(survival)
library(ncvreg)
library(dummies)
#install.packages(c("ncvreg", "dummies"))
```

Import data
```{r}
setwd("/Users/bryanlin/Desktop/zzzzzzzz files/survival final project")
breast_ca <- readRDS(file = "breast_final.rds")
```


## Variable selection

variables:

- stage: cancer stage
- delta: death indicator
- SRV_TIME_MON = Survival months survival time
- SEX = Sex
- RAC_RECY = race
- MAR_STAT = marital status at dx
- BRST_SUB = BREAST SUBTYPE
- AGE_DX = age at diagnosis
- YR_BRTH = year of birth
- MALIGCOUNT = Total number of in situ/malignant tumors for patient       
- BENBORDCOUNT = Total number of benign/borderline tumors for patient
- Age (year of end of study - YR_BRTH)

```{r}
source("Rfunction.R")
```

The design matrix of predictor values
```{r}
X <- breast_ca %>% 
  select(-SRV_TIME_MON, -delta)
X <- as.data.frame(X)
#get dummy variables
X <- dummy.data.frame(X, names = c('stage', 'SEX', 'RAC_RECY', 'MAR_STAT', 'BRST_SUB'))
X <- select(X, -stage0, -RAC_RECY1, -SEX1, -BRST_SUB1, -MAR_STAT1)
colnames(X)[colnames(X)=="SEX2"] <- "SEX"
```

Lasso
```{r}
n <- 300
lasso.fit <- ncvsurv(X, Surv(breast_ca$SRV_TIME_MON, breast_ca$delta), penalty = "lasso", nlambda = 25)
lasso.bic <- AIC(lasso.fit, k = log(n))
lasso.est <- lasso.fit$beta[, which.min(lasso.bic)]
```

SCAD
```{r}
scad.fit <- ncvsurv(X, Surv(breast_ca$SRV_TIME_MON, breast_ca$delta), penalty = "SCAD", nlambda = 25) 
scad.bic <- AIC(scad.fit, k = log(n))
scad.est <- scad.fit$beta[, which.min(scad.bic)]
```

MCP
```{r}
mcp.fit <- ncvsurv(X, Surv(breast_ca$SRV_TIME_MON, breast_ca$delta), penalty = "MCP", nlambda = 25)
mcp.bic <- AIC(mcp.fit, k = log(n))
mcp.est <- mcp.fit$beta[, which.min(mcp.bic)]
```

Results
```{r}
beta.est <- round(cbind(lasso.est, scad.est, mcp.est), 4)
beta.est
```


Fit KM plots
```{r}
library("survival")

#Define variables
time <- na.omit(breast_ca$SRV_TIME_MON)
delta <- na.omit(breast_ca$delta)
#as.numeric(breast_ca$SURV_TIME_MON))


fit <- survfit(Surv(time, delta) ~ 1)
summary(fit)

plot(fit, main = 'Product-Limit Survival Estimates', 
     ylab = 'Survival Probability', xlab = 'Time (Months)', 
     ylim = c(0.95, 1))



```


#Make KM plot for male vs. female
```{r}
#subset males; N = 68 males
males <- subset(breast_ca, breast_ca$SEX == 1) 
#count(males)
time_males <- males$SRV_TIME_MON
delta_males <- males$delta


#subset females; N = 9,948 females
females <- subset(breast_ca, breast_ca$SEX == 2)
#count(females)
time_females <- females$SRV_TIME_MON
delta_females <- females$delta



#gender indicator
M <- rep(1, 68) #68 males
FM <- rep(2, 9948) #9948 females


#fit model for gender
fit_gender <- survfit(Surv(c(time_males, time_females), c(delta_males, delta_females))
                      ~ c(M, FM))
summary(fit_gender)

#Make plot for gender
plot(fit_gender, lty = c(2, 1), col = c("blue", "red"), 
     main = 'Estimated Survival Functions for Breast Cancer \n by Gender',
     ylab = 'Estimted Survival Function', xlab = 'Time Months', ylim = c(0.95, 1))
legend('topright', c('Male', 'Female'), col = c('blue', 'red'), lty = 2:1, bty = 'n')
```

                      

#Make plot for different cancer stages
```{r}
#subset stage 1; N = 4822 patients
stg1 <- subset(breast_ca, breast_ca$stage == '1') 
count(stg1)
time_stg1 <- stg1$SRV_TIME_MON
delta_stg1 <- stg1$delta


#subset stage 2; N = 3,382 patients
stg2 <- subset(breast_ca, breast_ca$stage == '2')
count(stg2)
time_stg2 <- stg2$SRV_TIME_MON
delta_stg2 <- stg2$delta


#subset stage 3; N = 1,124 patients
stg3 <- subset(breast_ca, breast_ca$stage == '3')
count(stg3)
time_stg3 <- stg3$SRV_TIME_MON
delta_stg3 <- stg3$delta


#subset stage 4; N = 473 patients
stg4 <- subset(breast_ca, breast_ca$stage == '4')
count(stg4)
time_stg4 <- stg4$SRV_TIME_MON
delta_stg4 <- stg4$delta



#cancer stage indicator
st1 <- rep(1, 4822) #4822 stg1 patients
st2 <- rep(2, 3382) #3382 stg2 patients
st3 <- rep(3, 1124) #1124 stg3 patients
st4 <- rep(4, 473) #473 stg4 patients


#fit model for gender
fit_stage <- survfit(Surv(c(time_stg1, time_stg2, time_stg3, time_stg4), c(delta_stg1, delta_stg2, delta_stg3, delta_stg4))
                      ~ c(st1, st2, st3, st4))
summary(fit_stage)

#Make plot for cancer stage
plot(fit_stage, lty = c(1, 2, 3, 4), col = c("blue", "red", "green", "brown"), 
     main = 'Estimated Survival Functions for \n Breast Cancer Stages',
     ylab = 'Estimted Survival Function', xlab = 'Time (Months)', ylim = c(0.6, 1))
legend('bottomleft', c('Stage 1', 'Stage 2', 'Stage 3', 'Stage 4'), col = c('blue', 'red', 'green', 'brown'), lty = 1:4, bty = 'n')


```

#Make plot for different races (White, Black, American Indian or Alaska Native, Asian or Pacific Islander)


```{r}
#subset White ; N = 8505 patients
white <- subset(breast_ca, breast_ca$RAC_RECY == '1') 
count(white)
time_white <- white$SRV_TIME_MON
delta_white <- white$delta


#subset Black; N = 479 patients
black <- subset(breast_ca, breast_ca$RAC_RECY == '2')
count(black)
time_black <- black$SRV_TIME_MON
delta_black <- black$delta


#subset Indian or Alaska; N = 81 patients
indian <- subset(breast_ca, breast_ca$RAC_RECY == '3')
count(indian)
time_indian <- indian$SRV_TIME_MON
delta_indian <- indian$delta

#subset Asian; N = 951 patients
asian <- subset(breast_ca, breast_ca$RAC_RECY == '4')
count(asian)
time_asian <- asian$SRV_TIME_MON
delta_asian <- asian$delta


#cancer stage indicator
race1 <- rep(1, 8505) #8505 White patients
race2 <- rep(2, 479) #479 Black patients
race3 <- rep(3, 81) #81 Indian patients
race4 <- rep(4, 951) #951 Asian patients


#fit model for gender
fit_race <- survfit(Surv(c(time_white, time_black, time_indian, time_asian), c(delta_white, delta_black, delta_indian, delta_asian)) ~ c(race1, race2, race3, race4))
summary(fit_race)

#Make plot for cancer stage
plot(fit_race, lty = c(1, 2, 3, 4), col = c("blue", "red", "green", "brown"), 
     main = 'Estimated Survival Functions for \n Races',
     ylab = 'Estimted Survival Function', xlab = 'Time (Months)', ylim = c(0.90, 1))
legend('bottomleft', c('White', 'Black', 'Indian', 'Asian'), col = c('blue', 'red', 'green', 'brown'), lty = 1:4, bty = 'n')


```

